Deep Learning

### 2015-01-05 deepreview.tex
reformatted deepreview.tex into 80-line paragraphs to make it easier to see changes with diff
small changes as discussed, LSTM and beating previous methods, updated bib file

Deep learning is currently making major advances in solving problems that have resisted the best attempts of the
Artificial Intelligence community for many years. Deep learning is very good at discovering intricate structure in
large, high-dimensional datasets and is therefore widely applicable in all of the sciences. For example, it beats all
previous machine learning techniques at predicting the activity of potential drug molecules~\cite{Dahl}, detecting new
fundamental particles~\cite{??}, classifying galaxies~\cite{??}, or predicting the effects of mutations in non-coding
DNA on gene expression and disease~\cite{Frey}.

So far, most applications of deep learning use feed-forward neural networks, but for tasks that involve sequential
inputs it is often better to use recurrent neural networks. These are given one term from an input sequence at each time
step and their recurrent internal ``hidden'' state depends on the whole history of the input sequence. Recurrent
networks are very good at predicting the next character in a string\cite{Sutskever} or the next word in a
sequence\cite{Mikolov} but they can also be used for more complex tasks.  After reading an English sentence one word at
a time, for example, an English``encoder'' network can be trained so that the final state vector of its hidden neurons
is a good representation of the thought expressed by the sentence.  This thought vector can then be used as the initial
hidden state of a trained French ``decoder'' network which outputs a probability distribution for the first word of the
French translation. If a particuar first word is chosen and provided as input to the decoder network it will then output
a probability distribution for the second word of the translation and so on until a full stop is chosen~\cite{Vinyals, Montreal}.

@@

The obvious weakness of gradient descent is that it can get trapped in poor local minima, but in high-dimensional
weight-spaces this is much less common than getting stuck close to saddle points where the error surface curves upwards
in most directions but downwards in a few. Recent empirical\cite{Bengio} and theoretical results\cite{Yann} strongly
suggest that in deep networks almost all of these saddle points give very similar performance so it does not matter
which one we find.  This goes a long way towards explaining why stochastic gradient descent works so well for deep
neural networks.

Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by the introduction of an unsupervised learning
procedure that could create layers of feature detectors without requiring labeled data. The objective in learning each
layer of feature detectors was to be able to reconstruct or model the activities of feature detectors (or raw inputs) in the
layer below.  By ``pre-training''  several layers of progressively more complex feature detectors using this reconstruction
objective, the weights of a deep network could be initialized to sensible values.  A final layer of output units could
then be added to the top of the network and the whole deep system could be fine-tuned using standard
backpropagation\cite{Science}. This worked remarkably well for recognizing handwritten digits especially when the amount
of labeled data was very limited\cite{BengioNips06}.

The first major application of this pre-training approach was in speech recognition and it was made possible by the
advent of fast Graphics Processing Units (GPUs) that were convenient to program.  In 2009, the approach was used to map
short temporal windows of coefficients extracted from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.  It achieved record-breaking results on a
standard speech recognition benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly developed to give
record-breaking results on a large vocabulary task\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\cite{SPS} and were already being deployed in Android phones.  For smaller
datasets, unsupervised pre-training leads to significantly better generalization, but once deep
learning had been rehabilitated, it turned out that large labeled datasets did not need the pre-training stage, provided the
scales of the initial random weights were carefully chosen to avoid very large or very small initial gradients.

Interest in recurrent networks (which are deep in time) was revived by ``Long Short Term Memory'' networks (LSTMs) that
use special hidden units whose natural behaviour is to remember inputs for a long time\cite{Hochreiter}.  A
linear ``memory cell'' interacts with other units via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input vector has little effect on the state of a
memory cell if the cell's input gate has a low activation so memory cells can learn to selectively ignore the input.  Similarly,
a memory cell has little effect on the output if its output gate has low activation. Each memory cell also has a
connection to itself at the next time-step that has a weight of 1 so it tends to copy its own real-valued state, but
this connection is gated by a ``keep gate'' so the stored value is quickly eroded unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by weighted inputs that they receive from the input
vectors and all of the memory cells.

When LSTMs were first described in English\cite{HochSchmid} their complexity was off-putting and, despite their
impressive ability to learn very long-range temporal structure on toy problems, they were largely ignored until they
achieved excellent results in a competition to read cursive hand-writing\cite{Graves}. They have subsequently proved to
be a very effective for tasks that have sequential input.  They are now better than deep feed-forward networks for
speech recognition\cite{gravestimit} and unlike feed-forward nets they can implement an entire speech recognition system
that goes all the way from coefficents derived from the soundwave to the sequence of characters in the transcription~\cite{graves}.
They are also what is currently used for the encoder and decoder networks that perform so well at machine translation.

It is not yet clear which aspects of the complicated architecture of LSTMs are really necessary. Memory cells are a good
way of allowing the backpropagated error derivatives to neither explode nor vanish, but they are not the only way.  For
example, recurrent nets with rectified linear hidden units that use the non-linear function $y=max(0,x)$ are much
simpler than LSTMs and if the recurrent weight matrix is initialized to be close to the identity matrix they work about
as well as LSTMs both for toy problems involving very long-range temporal structure and for real tasks like predicting
the next word in text\cite{LeHinton}.

>>>>

Deep learning is currently making major advances in solving problems that
have resisted the best attempts of the Artificial Intelligence community
for many years. Deep learning is very good at discovering intricate
structure in large, high-dimensional datasets and is therefore widely
applicable in all of the sciences. For example, it has beaten previous
machine learning techniques at predicting the activity of potential drug
molecules~\cite{Dahl}, detecting new fundamental particles~\cite{??},
classifying galaxies~\cite{??}, or predicting the effects of mutations in
non-coding DNA on gene expression and disease~\cite{Frey}.

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs it is often better
to use recurrent neural networks. These are given one element from an input % YB term-->element
sequence at each time step and their recurrent internal ``hidden'' state
depends on the whole history of the input sequence. Recurrent networks are
very good at predicting the next character in a string\cite{Sutskever} or
the next word in a sequence\cite{Mikolov} but they can also be used for
more complex tasks.  After reading an English sentence one word at a time,
for example, an English``encoder'' network can be trained so that the final
state vector of its hidden neurons is a good representation of the thought
expressed by the sentence.  This thought vector can then be used as the
initial hidden state of a trained French ``decoder'' network which outputs
a probability distribution for the first word of the French translation. If
a particular first word is chosen (e.g., according to that probability)
and provided as input to the decoder
network it will then output a probability distribution for the second word
of the translation and so on until a full stop is chosen~\cite{Vinyals,
  Montreal}. 
%YB added a clarification
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\cite{Dauphin-et-al-NIPS2014-small,Yann} strongly suggest
that in deep networks almost all of these local minima give very similar
performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by
the introduction of an unsupervised learning procedure that could create
layers of feature detectors without requiring labeled data. The objective
in learning each layer of feature detectors was to be able to reconstruct
or model the activities of feature detectors (or raw inputs) in the layer
below.  By ``pre-training'' several layers of progressively more complex
feature detectors using this reconstruction objective, the weights of a
deep network could be initialized to sensible values.  A final layer of
output units could then be added to the top of the network and the whole
deep system could be fine-tuned using standard
backpropagation\cite{Science}. This worked remarkably well for recognizing
handwritten digits especially when the amount of labeled data was very
limited\cite{Bengio-nips-2006-small}.

The first major application of this pre-training approach was in speech
recognition and it was made possible by the advent of fast Graphics
Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
In 2009, the approach was used to map short temporal windows of coefficients extracted
from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.
It achieved record-breaking results on a standard speech recognition
benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly
developed to give record-breaking results on a large vocabulary
task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\cite{SPS} and were already
being deployed in Android phones.  For smaller datasets, unsupervised
pre-training leads to significantly better generalization~\citep{Erhan+al-2010-small}, 
but once deep
learning had been rehabilitated, it turned out that large labeled datasets
did not need the pre-training stage, provided the scales of the initial
random weights were carefully chosen to avoid very large or very small
initial gradients.

Interest in recurrent networks (which are deep in time) was revived by
``Long Short Term Memory'' networks (LSTMs) that use special hidden units
whose natural behaviour is to remember inputs for a long
time~\cite{Hochreiter+Schmidhuber-1997}.  
A linear ``memory cell'' interacts with other units
via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input
vector has little effect on the state of a memory cell if the cell's input
gate has a low activation so memory cells can learn to selectively ignore
the input.  Similarly, a memory cell has little effect on the output if its
output gate has low activation. Each memory cell also has a connection to
itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
own real-valued state, but this self-connection is gated by a ``keep gate'' so
the stored value leaks quickly unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by
weighted inputs that they receive from the input vectors and all of the
memory cells.

When LSTMs were first described in
English~\cite{Hochreiter+Schmidhuber-1997} their complexity was off-putting
and, despite their impressive ability to learn very long-range temporal
structure on toy problems, they were largely ignored until they achieved
excellent results in a competition to read cursive
hand-writing~\cite{Graves}. They have subsequently proved to be very
effective for tasks that have sequential input.  They are now better than
deep feed-forward networks for speech recognition~\cite{gravestimit} and
unlike feed-forward nets they can implement an entire speech recognition
system that goes all the way from coefficents derived from the soundwave to
the sequence of characters in the transcription~\cite{gravestimit}.  They
are also what is currently used for the encoder and decoder networks that
perform so well at machine
translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

Recent work~\citep{Chung-et-al-NIPSDL2014} suggests that not all the elements of the LSTM
architecture are required, the crucial element being the presence of
of the gate controlling the leaking rate (or time constant) of the cell.
%It is not yet clear which aspects of the complicated architecture of LSTMs
%are really necessary. Memory cells are a good way of allowing the
%backpropagated error derivatives to neither explode nor vanish, but they
%are not the only way.  For example, recurrent nets with rectified linear
%hidden units that use the non-linear function $y=max(0,x)$ are much simpler
%than LSTMs and if the recurrent weight matrix is initialized to be close to
%the identity matrix they work about as well as LSTMs both for toy problems
%involving very long-range temporal structure and for real tasks like
%predicting the next word in text\cite{LeHinton}.
